### 2022-03-25 ehoh2/ehoh2.tex
more

>>>>

  \section{Discussion and Related Work} %---0.5p
  \label{sec:ehoh2:discussion}

  \looseness=-1
  On the trajectory to \ehohii{}, we developed, together with colleagues, three
  superposition calculi:\ \lfsup{} for $\lambda$-free higher-order logic
  \cite{bbcw-21-lfho}, \lsup{} for a higher-order logic with
  $\lambda$-abstra\-ction but no Booleans \cite{bbtvw-21-sup-lam}, and \osup{}
  for full higher-order logic \cite{bbtvw-21-sup-lam}. These milestones allowed
  us to carefully estimate how the increased reasoning capabilities of each
  calculus influence its performance. We also developed a complete and efficient
  unification algorithm (Chapter \ref{ch:unif}) which can easily be customized
  to trade bits of its completeness for efficiency. Practically oriented work
  described in Chapters \ref{ch:bools} and \ref{ch:ho-techniques},
  helped us gauge how much reasoning power we lose by using more intuitive and
  more easily controllable incomplete rules. Both the theoretical and the
  practical lines of our work met in this chapter, resulting in a prover that
  outperforms the competition on the benchmarks from proof assistants.

  Extending first-order provers with higher-order reasoning capabilities has been
  attempted by other researchers as well.
  Barbosa et al.\ extended
  the SMT solvers CVC4 (now cvc5) and veriT to higher-order logic in an incomplete
  way \cite{brotb-19-ho-smt}.
  Bhayat and Reger first extended Vampire to
  higher-order logic using combinatory unification \cite{br-19-restricted-unif},
  an incomplete approach, before they designed and implemented a complete
  higher-order superposition calculus based on \textsf{SKBCI}-com\-binators \cite{br-20-full-sup-w-combs}.
  The advantage is that combinators can be supported as a thin layer on top of
  $\lambda$-free terms.
  This calculus is also implemented in Zipperposition.
  However, in unpublished experiments, we found that $\lambda$-superposition
  performs substantially better, corroborating the CASC results, so we decided to
  make a more profound change to Ehoh and implement $\lambda$-superposition.
  \pagebreak[2]

  Possibly the only actively maintained higher-order prover built from the bottom
  up as a higher-order prover is Leo-III \cite{sb-21-leo3}, as the long-time
  winner of CASC higher-order division, Satallax \cite{cb-12-satallax}, is no
  longer maintained. We give a further overview of other traditional higher-order
  provers and the calculi they are based in Sect.~\ref{sec:ehoh:discussion-and-related-work}.

  \section{Conclusion} %---0.5p
  \label{sec:ehoh2:conclusion}

  In 2019, the reviewers of our paper introducing Ehoh \cite{vbss-19-ehoh1} were
  skeptical that extending it with support for full higher-order logic would be
  feasible. One of them wrote:
  %
  \begin{quote}
  A potential criticism could be that this step from E to Ehoh is just extending
  FOL by those aspects of HOL that are easily in reach with rather straightforward
  extensions (none of the extensions is indeed very complicated), and that the
  difficult challenges of fully supporting HOL have yet to be confronted.
  \end{quote}
  %
  We ended up addressing the theoretical ``difficult challenges'' in other work
  with colleagues. In this chapter, we faced the practical challenges pertaining to the extension
  of Ehoh's data structures and algorithms to support full higher-order logic and
  demonstrated that such an extension is possible. \NumberOK{Our evaluation} shows
  that this extension makes \ehohii{} the best higher-order prover on
  benchmarks coming from interactive theorem proving practice, which has been our goal since 2016. \ehohii{}'s
  performance lags slightly behind Zipperposition on TPTP problems, possibly because Zipperposition
  does not assume a clausal structure and can perform subtle formula-level
  inferences. In the future, we plan to implement the same features in \ehohii{}.
  We have also only started tuning \ehohii{}'s heuristics on higher-order
  problems.
